analysis_uown

### 2021-11-15 code/analysis_code/analysis_uown.Rmd
Supplementary_Data - CC

---
title: "analysis_uown"
author: "C. Coleman"
date: "10/27/2021"
output: html_document
---

@@

_Therefore, the following constitutes the statistical analysis protocol that will be conducted in this markdown:_
</br>
#### PART 1: Outliers in stream health indicator (Biological Score, E. Coli (cfu), Conductivity, NO3 (mg/L), and Turbidity) linear relationships.
_1. Assess each linear relationship previously performed for outliers in the data;_
_2. Determine if it is appropriate to remove outliers;_
_3. Rerun linear regressions for the different stream health indicator relationships and assess if more significant relationships exist._
</br>
#### Part 2: Linear Model Fit Analysis of MIDO Data
_1. Conduct test/train data splitting for MIDO Data._
_2. Conduct linear modeling for relationships between stream health indicator predictors (Biological Score, E. Coli (cfu), Conductivity, NO3 (mg/L), and Turbidity) in the form of a workflow for the Test and Train MIDO data._
_3. Predict values for Test and Train MIDO data._
_4. Assess how well the MIDO linear model suite of relationships (Biological Score, E. Coli (cfu), Conductivity, NO3 (mg/L), and Turbidity) fir the test and train data using Root Mean Square Error (rmse). The result will measure how good of a fit the created models are to the data._

### Required packages
```{r}
library(tidyverse) #Working with multiple Tidy packages
library(tidymodels) #Building models
library(dplyr) #data manipulation
library(here) #setting pathways for saving files
library(rpart) #Model fitting
library(ranger) #Model fitting
library(glmnet) #Model fitting
```

@@

ecfu_con_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

@@

ecfu_con_aug %>%
  rmse(truth = e.coli.cfu, .pred)

@@

ecfu_turb_aug %>%
  rmse(truth = e.coli.cfu, .pred)
```
</br>

#####Analysis here#####

@@

ecfu_con_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()

@@

ecfu_turb_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()
```
#####Analysis here#####

@@

ecfu_con_aug_test %>%
  rmse(truth = e.coli.cfu, .pred)

@@

#### Save significant Test and Train linear models to the results folder.
```{r}
#Create a location and save linear model for E. coli (cfu) in the MIDO train data.
MIDO_ECFU_Con_fit_location <- here::here("results","MIDO_ECFU_Con_fit.rds")
saveRDS(ecfu_con_fit, file = MIDO_ECFU_Con_fit_location)

#Create a location and save linear model for E. coli (cfu) in the MIDO test data.
MIDO_ECFU_Con_fit_test_location <- here::here("results","MIDO_ECFU_Con_fit_test.rds")
saveRDS(ecfu_con_fit_test, file = MIDO_ECFU_Con_fit_test_location)
```

@@

#### Save significant Test and Train RMSE to the results folder.
```{r}
#Create a location and save RMSE for E. coli (cfu) in the MIDO train data.
MIDO_ECFU_Con_RMSE_location <- here::here("results","MIDO_ECFU_Con_RMSE.rds")
saveRDS(ecfu_con_aug, file = MIDO_ECFU_Con_RMSE_location)

#Create a location and save linear RMSE for E. coli (cfu) in the MIDO test data.
MIDO_ECFU_Con_RMSE_test_location <- here::here("results","MIDO_ECFU_Con_RMSE_test.rds")
saveRDS(ecfu_con_aug_test, file = MIDO_ECFU_Con_RMSE_test_location)
```

@@

@@

_Showing and selecting best performing Models_
```{r}
#Showing best performing LASSO models
lasso_resample %>%
  show_best()

#Selects best performing model
best_lasso <- lasso_resample %>%
  select_best()
```

@@

lasso_residuals
```
_model predictions from tuned model vs actual outcomes_
```{r}
lasso_pred_plot <- ggplot(lasso_residuals, 
                          aes(x = biological_score, 
                              y = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Actual: LASSO", 
       x = "Biological Score Outcome", 
       y = "Biological Score Prediction")
lasso_pred_plot
```
_plot residuals vs predictions_
```{r}
lasso_residual_plot <- ggplot(lasso_residuals, 
                              aes(y = .resid, 
                                  x = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Residuals: LASSO", 
       x = "Biological Score Prediction", 
       y = "Residuals")
plot(lasso_residual_plot) #view plot
```

@@

@@

_Showing and selecting best performing Models_
```{r}
#Showing best performing LASSO models
lasso_resample_EC %>%
  show_best()

#Selects best performing model
best_lasso_EC <- lasso_resample_EC %>%
  select_best()
```

@@

lasso_residuals_EC
```
_model predictions from tuned model vs actual outcomes_
```{r}
lasso_pred_plot_EC <- ggplot(lasso_residuals_EC, 
                          aes(x = e.coli.cfu, 
                              y = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Actual: LASSO", 
       x = "E. coli cfu Outcome", 
       y = "E. coli cfu Prediction")
lasso_pred_plot_EC
```
_plot residuals vs predictions_
```{r}
lasso_residual_plot_EC <- ggplot(lasso_residuals_EC, 
                              aes(y = .resid, 
                                  x = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Residuals: LASSO", 
       x = "E. coli cfu Prediction", 
       y = "Residuals")
plot(lasso_residual_plot_EC) #view plot
```

>>>>

---
title: "analysis_uown"
author: "C. Coleman"
date: "10/27/2021"
output:
  word_document: default
  html_document: default
---

@@

_Therefore, the following constitutes the statistical analysis protocol that will be conducted in this markdown:_
</br>
#### PART 1: Outliers in stream health indicator (Biological Score, E. Coli (cfu), Conductivity, NO3 (mg/L), and Turbidity) linear relationships.
_1. Assess each linear relationship previously performed for outliers in the data;_
_2. Determine if it is appropriate to remove outliers;_
_3. Rerun linear regressions for the different stream health indicator relationships and assess if more significant relationships exist._

#### PART 2: Linear Model Fit Analysis of MIDO Data
_1. Conduct test/train data splitting for MIDO Data._
_2. Conduct linear modeling for relationships between stream health indicator predictors (Biological Score, E. Coli (cfu), Conductivity, NO3 (mg/L), and Turbidity) in the form of a workflow for the Test and Train MIDO data._
_3. Predict values for Test and Train MIDO data._
_4. Assess how well the MIDO linear model suite of relationships (Biological Score, E. Coli (cfu), Conductivity, NO3 (mg/L), and Turbidity) fir the test and train data using Root Mean Square Error (rmse). The result will measure how good of a fit the created models are to the data._

#### PART 3: LASSO Modeling MIDO Data
_Conduct LASSO modeling of linear regression between Biological Score and all other variables and E. coli CFU and all other variables. Cross validation well be done by calculating RMSE for all models and compared to the RMSE for linear regression modeling of a Null model. Additionally, The best model will be selected using the R function select_best(), for which residuals will be calculated between predicted and actual data._

### Required packages
```{r}
library(tidyverse) #Working with multiple Tidy packages
library(tidymodels) #Building models
library(dplyr) #data manipulation
library(here) #setting pathways for saving files
library(rpart) #Model fitting
library(ranger) #Model fitting
library(glmnet) #Model fitting
library(knitr) #saving tables
```

@@

Ecoli_con_lm <- ecfu_con_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

@@

Ecoli_con_RMSE <- ecfu_con_aug %>%
  rmse(truth = e.coli.cfu, .pred)

@@

ecfu_turb_aug %>%
  rmse(truth = e.coli.cfu, .pred)
```

@@

Ecoli_con_test_lm <- ecfu_con_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()

@@

ecfu_turb_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

@@

Ecoli_con_RMSE_test <- ecfu_con_aug_test %>%
  rmse(truth = e.coli.cfu, .pred)

@@

#### Save significant Test and Train linear models to the results folder.
```{r}
#Create a table for E. coli (cfu) as a function of Conductivity Linear Fit Model in the MIDO train data.
MIDO_ECFU_Con_fit <- kable(as.tibble(Ecoli_con_lm),
      caption = "Linear Fit Model: E. Coli CFU predicted by Conductivity in the Trained MIDO Data Set")

MIDO_ECFU_Con_fit

#Create a table for E. coli (cfu) as a function of Conductivity Linear Fit Model in the MIDO test data.

MIDO_ECFU_Con_fit_test <- kable(as.tibble(Ecoli_con_test_lm),
      caption = "Linear Fit Model: E. Coli CFU predicted by Conductivity in the Test MIDO Data Set")

MIDO_ECFU_Con_fit_test
```

@@

#### Save significant Test and Train RMSE to the results folder.
```{r}
#Create a table for E. coli (cfu) as a function of Conductivity Linear Fit Model RMSE in the MIDO train data.
MIDO_ECFU_Con_RMSE <- kable(as.tibble(Ecoli_con_RMSE),
      caption = "Root Mean Square Error: E. Coli CFU predicted by Conductivity in the Trained MIDO Data Set")

MIDO_ECFU_Con_RMSE

#Create a table for E. coli (cfu) as a function of Conductivity Linear Fit Model RMSE in the MIDO test data.

MIDO_ECFU_Con_RMSE_test <- kable(as.tibble(Ecoli_con_RMSE),
      caption = "Root Mean Square Error: E. Coli CFU predicted by Conductivity in the Test MIDO Data Set")

MIDO_ECFU_Con_RMSE_test
```

@@

Null_Met <- kable(Null_Met,
      caption = "Root Mean Squared Error and Standard Deviation for Biological Score Null Linear Model")

@@

_Showing and selecting best performing Models_
```{r}
#Showing best performing LASSO models
top_five_bs <- lasso_resample %>%
  show_best()

#Selects best performing model
best_lasso <- lasso_resample %>%
  select_best()

Best_LASSO_BS <- kable(top_five_bs,
      caption = "Top 5 best performing LASSO Models for Biological Score as the Outcome")

Best_LASSO_BS
```

@@

lasso_residuals
```
_model predictions from tuned model vs actual outcomes_
```{r}
lasso_pred_plot <- ggplot(lasso_residuals, 
                          aes(x = biological_score, 
                              y = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Actual: LASSO", 
       x = "Biological Score Outcome", 
       y = "Biological Score Prediction")
lasso_pred_plot

#save plot
lasso_pred_plot_BS_location <- here::here("results", "lasso_prediction_plot_BS.png")

ggsave(lasso_pred_plot_BS_location)
```

_plot residuals vs predictions_
```{r}
lasso_residual_plot <- ggplot(lasso_residuals, 
                              aes(y = .resid, 
                                  x = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Residuals: LASSO", 
       x = "Biological Score Prediction", 
       y = "Residuals")
plot(lasso_residual_plot) #view plot
```

@@

Null_Met_EC <- kable(Null_Met_EC,
      caption = "Root Mean Squared Error and Standard Deviation for E. coli cfu Null Linear Model")

@@

_Showing and selecting best performing Models_
```{r}
#Showing best performing LASSO models
top_five_ec <- lasso_resample_EC %>%
  show_best()

#Selects best performing model
best_lasso_EC <- lasso_resample_EC %>%
  select_best()

Best_LASSO_EC <- kable(top_five_ec,
      caption = "Top 5 best performing LASSO Models for E. coli cfu as the Outcome")

Best_LASSO_EC
```

@@

lasso_residuals_EC
```
_model predictions from tuned model vs actual outcomes_
```{r}
lasso_pred_plot_EC <- ggplot(lasso_residuals_EC, 
                          aes(x = e.coli.cfu, 
                              y = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Actual: LASSO", 
       x = "E. coli cfu Outcome", 
       y = "E. coli cfu Prediction")
lasso_pred_plot_EC

#save plot
lasso_pred_plot_EC_location <- here::here("results", "lasso_prediction_plot_EC.png")

ggsave(lasso_pred_plot_EC_location)
```
_plot residuals vs predictions_
```{r}
lasso_residual_plot_EC <- ggplot(lasso_residuals_EC, 
                              aes(y = .resid, 
                                  x = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Residuals: LASSO", 
       x = "E. coli cfu Prediction", 
       y = "Residuals")
plot(lasso_residual_plot_EC) #view plot
```
